# add_images

Add images to an existing image collection



```{r include=FALSE}
library(gdalcubes)
```

## Description

This function adds provided files or GDAL dataset identifiers and to an existing image collection by extracting datetime, image identifiers, and band information according to the collection's format.


## Usage


```r
add_images(
  image_collection,
  files,
  unroll_archives = TRUE,
  out_file = "",
  quiet = FALSE
)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| image_collection | image_collection object or path to an existing collection file |
| files | character vector with paths to image files on disk or any GDAL dataset identifiers (including virtual file systems and higher level drivers or GDAL subdatasets) |
| unroll_archives | automatically convert .zip, .tar archives and .gz compressed files to GDAL virtual file system dataset identifiers (e.g. by prepending /vsizip/) and add contained files to the list of considered files |
| out_file | path to output file, an empty string (the default) will update the collection in-place, whereas images will be added to a new copy of the image collection at the given location otherwise. |
| quiet | logical; if TRUE, do not print resulting image collection if return value is not assigned to a variable |


## Value

image collection proxy object, which can be used to create a data cube using [`raster_cube`](raster_cube.Rmd)


## Examples


```{r}
L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
L8_col = create_image_collection(L8_files[1:12], "L8_L1TP") 
add_images(L8_col, L8_files[13:24])
```


