# apply_pixel.cube

Apply arithmetic expressions over all pixels of a data cube



```{r include=FALSE}
library(gdalcubes)
```

## Description

Create a proxy data cube, which applies arithmetic expressions over all pixels of a data cube. Expressions may access band values by name.


## Usage


```r
apply_pixel.cube(x, expr, names = NULL, keep_bands = FALSE, ..., FUN)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| x | source data cube |
| expr | character vector with one or more arithmetic expressions (see Details) |
| names | optional character vector with the same length as expr to specify band names for the output cube |
| keep_bands | logical; keep bands of input data cube, defaults to FALSE, i.e. original bands will be dropped |
| ... | not used |
| FUN | user-defined R function that is applied on all pixels (see Details) |


## Details

The function can either apply simple arithmetic C expressions given as a character vector (expr argument), or apply a custom R reducer function if FUN is provided.

In the former case, gdalcubes uses the [tinyexpr library](https://github.com/codeplea/tinyexpr) to evaluate expressions in C / C++, you can look at the [library documentation](https://github.com/codeplea/tinyexpr#functions-supported)
to see what kind of expressions you can execute. Pixel band values can be accessed by name.

FUN receives values of the bands from one pixel as a (named) vector and should return a numeric vector with identical length for all pixels. Elements of the
result vectors will be interpreted as bands in the result data cube.


## Value

a proxy data cube object


## Note

This function returns a proxy object, i.e., it will not start any computations besides deriving the shape of the result.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

# 1. Apply a C expression
L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-04", t1="2018-06"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
L8.cube = raster_cube(L8.col, v) 
L8.cube = select_bands(L8.cube, c("B04", "B05")) 
L8.ndvi = apply_pixel(L8.cube, "(B05-B04)/(B05+B04)", "NDVI") 
L8.ndvi


plot(L8.ndvi)


# 2. Apply a user defined R function
L8.ndvi.noisy = apply_pixel(L8.cube, names="NDVI_noisy", 
   FUN=function(x) {
       rnorm(1, 0, 0.1) + (x["B05"]-x["B04"])/(x["B05"]+x["B04"])
   })
L8.ndvi.noisy


plot(L8.ndvi.noisy)

 
```


