# as_json

Query data cube properties



```{r include=FALSE}
library(gdalcubes)
```

## Description

gdalcubes internally uses a graph to serialize data cubes (including chained operations on cubes). This function derives a JSON
representation, which can be used to save data cube objects without pixel data to disk.


## Usage


```r
as_json(obj, file = NULL)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| obj | a data cube proxy object (class cube) |
| file | optional output file |


## Value

If file is NULL, the function returns a JSON string representing a graph that can be used to recreate the same
chain of gdalcubes operations even in a different R sessions.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-04", t1="2018-04"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
cat(as_json(select_bands(raster_cube(L8.col, v), c("B04", "B05"))))
```


