# chunk_apply

Apply an R function on chunks of a data cube



```{r include=FALSE}
library(gdalcubes)
```

## Description

Apply an R function on chunks of a data cube


## Usage


```r
chunk_apply(cube, f)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| cube | source data cube |
| f | R function to apply over all chunks |


## Details

This function internally creates a gdalcubes stream data cube, which streams
data of a chunk to a new R process. For reading data, the function typically 
calls `x <- read_chunk_as_array()` which then results in a 4 dimensional (band, time, y, x) array.
Similarly `write_chunk_from_array(x)` will write a result array as a chunk in the resulting data cube.
The chunk size of the input cube is important to control how the function will be exposed to the data cube. For example,
if you want to apply an R function over complete pixel time series, you must define the chunk size argument in [`raster_cube`](raster_cube.Rmd)
to make sure that chunk contain the correct parts of the data.


## Value

a proxy data cube object


## Note

This function returns a proxy object, i.e., it will not start any computations besides deriving the shape of the result.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4,
                          bottom=4345299, top=4744931, t0="2018-01", t1="2018-12"),
                          srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
L8.cube = raster_cube(L8.col, v)
L8.cube = select_bands(L8.cube, c("B04", "B05"))
f <- function() {
  x <- read_chunk_as_array()
  out <- reduce_time(x, function(x) {
    cor(x[1,], x[2,], use="na.or.complete", method = "kendall")
  }) 
  write_chunk_from_array(out)
}
L8.cor = chunk_apply(L8.cube, f)
```


