# collection_formats

List predefined image collection formats



```{r include=FALSE}
library(gdalcubes)
```

## Description

gdalcubes comes with some predefined collection formats e.g. to scan Sentinel 2 data. This function lists available formats  including brief descriptions.


## Usage


```r
collection_formats(print = TRUE)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| print | logical; should available formats and their descriptions be printed nicely, defaults to TRUE |


## Details

Image collection formats define how individual files / GDAL datasets relate to an image collection, i.e., 
which bands they contain, to which image they belong, and how to derive aquisition date/time.
They are described as a set of regular expressions in a JSON file and used by gdalcubes to extract this information 
from the paths and/or filenames.


## Value

data.frame with columns `name` and `description` where the former describes the unique identifier that can be used in `create_image_collection` and the
latter gives a brief description of the format.


## Examples


```{r}
collection_formats()
```


