# image_mask

Create a mask for images in a raster data cube



```{r include=FALSE}
library(gdalcubes)
```

## Description

Create an image mask based on a band and provided values to filter pixels of images 
read by [`raster_cube`](raster_cube.Rmd)


## Usage


```r
image_mask(
  band,
  min = NULL,
  max = NULL,
  values = NULL,
  bits = NULL,
  invert = FALSE
)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| band | name of the mask band |
| min | minimum value, values between `min` and `max` will be masked |
| max | maximum value, values between `min` and `max` will be masked |
| values | numeric vector; specific values that will be masked. |
| bits | for bitmasks, extract the given bits (integer vector) with a bitwise AND before filtering the mask values, bit indexes are zero-based |
| invert | logical; invert mask |


## Details

Values of the selected mask band can be based on a range (by passing `min` and `max`) or on a set of values (by passing `values`). By default
pixels with mask values contained in the range or in the values are masked out, i.e. set to NA. Setting `invert = TRUE` will invert the masking behavior.
Passing `values` will override `min` and `max`.


## Note

Notice that masks are applied per image while reading images as a raster cube. They can be useful to eliminate e.g. cloudy pixels before applying the temporal aggregation to
merge multiple values for the same data cube pixel.


## Examples


```{r}
image_mask("SCL", values = c(3,8,9)) # Sentinel 2 L2A: mask cloud and cloud shadows
image_mask("BQA", bits=4, values=16) # Landsat 8: mask clouds
image_mask("B10", min = 8000, max=65000) 
```


