# image_collection

Load an existing image collection from a file



```{r include=FALSE}
library(gdalcubes)
```

## Description

This function will load an image collection from an SQLite file. Image collection files
index and reference existing imagery. To create a collection from files on disk,
use [`create_image_collection`](create_image_collection.Rmd).


## Usage


```r
image_collection(path)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| path | path to an existing image collection file |


## Value

an image collection proxy object, which can be used to create a data cube using [`raster_cube`](raster_cube.Rmd)


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
L8.col
```


