# json_cube

Read a data cube from a json description file



```{r include=FALSE}
library(gdalcubes)
```

## Description

Read a data cube from a json description file


## Usage


```r
json_cube(json, path = NULL)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| json | length-one character vector with a valid json data cube description |
| path | source data cube proxy object |


## Details

Data cubes can be stored as JSON description files. These files do not store any data but the recipe
how a data cube is constructed, i.e., the chain (or graph) of processes involved. 

Since data cube objects (as returned from [`raster_cube`](raster_cube.Rmd)) cannot be saved with normal R methods,
the combination of [`as_json`](as_json.Rmd) and [`json_cube`](json_cube.Rmd) provides a cheap way to save virtual 
data cube objects across several R sessions, as in the examples.


## Value

data cube proxy object


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-01", t1="2018-12"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
cube = raster_cube(L8.col, v) 

# save
fname = tempfile()
as_json(cube, fname)

# load
json_cube(path = fname)  
```


