# ncdf_cube

Read a data cube from an existing netCDF file



```{r include=FALSE}
library(gdalcubes)
```

## Description

Create a proxy data cube from a netCDF file that has been created using [`write_ncdf`](write_ncdf.Rmd).
This function does not read cubes from arbitrary netCDF files and can be used e.g., to load intermediate results and/or 
plotting existing netCDF cubes on disk without doing the data cube creation from image collections.


## Usage


```r
ncdf_cube(path, chunking = NULL, auto_unpack = TRUE)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| path | path to an existing netCDF file |
| chunking | custom chunk sizes to read form the netCDF file; defaults to using chunk sizes from the netCDF file |
| auto_unpack | logical; automatically apply offset and scale when reading data values |


## Value

a proxy data cube object


## Note

This function returns a proxy object, i.e., it will not start any computations besides deriving the shape of the result.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-04", t1="2018-06"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")


ncfile = write_ncdf(select_bands(raster_cube(L8.col, v), c("B02", "B03", "B04")))
ncdf_cube(ncfile)
                           
```


