# rename_bands

Rename bands of a data cube



```{r include=FALSE}
library(gdalcubes)
```

## Description

Create a proxy data cube, which renames specific bands of a data cube.


## Usage


```r
rename_bands(cube, ...)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| cube | source data cube |
| ... | named arguments with bands that will be renamed, see Details |


## Details

The result data cube always contains the same number of bands. No subsetting is 
done if only names for some of the bands are provided. In this case, only 
provided bands are renamed whereas other bands keep their original name. 
Variable arguments must be named by the old band name and the new names must be provided
as simple character values (see example).


## Value

proxy data cube object


## Note

This function returns a proxy object, i.e., it will not start any computations besides deriving the shape of the result.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-04", t1="2018-07"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
L8.cube = raster_cube(L8.col, v) 
L8.rgb = select_bands(L8.cube, c("B02", "B03", "B04"))
L8.rgb
L8.rgb = rename_bands(L8.cube, B02 = "blue", B03 = "green", B04 = "red")
L8.rgb
```


