# select_bands

Select bands of a data cube



```{r include=FALSE}
library(gdalcubes)
```

## Description

Create a proxy data cube, which selects specific bands of a data cube. The resulting cube
will drop any other bands.


## Usage


```r
select_bands(cube, bands)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| cube | source data cube |
| bands | character vector with band names |


## Value

proxy data cube object


## Note

This function returns a proxy object, i.e., it will not start any computations besides deriving the shape of the result.

For performance reasons, `select_bands` should always be called directly on a cube created with [`raster_cube`](raster_cube.Rmd) and 
drop all unneded bands. This allows to reduce RasterIO and warp operations in GDAL.


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
              bottom=4345299, top=4744931, t0="2018-04", t1="2018-07"),
              srs="EPSG:32618", nx = 497, ny=526, dt="P1M")
L8.cube = raster_cube(L8.col, v) 
L8.rgb = select_bands(L8.cube, c("B02", "B03", "B04"))
L8.rgb

plot(L8.rgb, rgb=3:1)
```


