# window_time

Apply a moving window operation over time



```{r include=FALSE}
library(gdalcubes)
```

## Description

This generic function applies a reducer function over a moving window over the time dimension of a data cube, an R array, or other classes if implemented.


## Usage


```r
window_time(x, ...)

```


## Arguments

| Argument    | Description                       |
|:------------|:----------------------------------|
| x | object to be reduced |
| ... | further arguments passed to specific implementations |


## Value

value and type depend on the class of x


## Examples


```{r}
# create image collection from example Landsat data only 
# if not already done in other examples
if (!file.exists(file.path(tempdir(), "L8.db"))) {
  L8_files <- list.files(system.file("L8NY18", package = "gdalcubes"),
                         ".TIF", recursive = TRUE, full.names = TRUE)
  create_image_collection(L8_files, "L8_L1TP", file.path(tempdir(), "L8.db"), quiet = TRUE) 
}

L8.col = image_collection(file.path(tempdir(), "L8.db"))
v = cube_view(extent=list(left=388941.2, right=766552.4, 
                          bottom=4345299, top=4744931, t0="2018-01", t1="2018-07"),
                          srs="EPSG:32618", nx = 400, dt="P1M")
L8.cube = raster_cube(L8.col, v) 
L8.nir = select_bands(L8.cube, c("B05"))
window_time(L8.nir, window = c(2,2), "min(B05)")  
window_time(L8.nir, kernel=c(-1,1), window=c(1,0))


plot(window_time(L8.nir, kernel=c(-1,1), window=c(1,0)), key.pos=1)
 
```


## See Also

[`window_time.cube`](window_time.cube.Rmd)


